Model Checking for Open Concurrent Systems

### 2018-09-10 2018-ModelCheckingPNets/checking.tex
new version with figure problem

%******Figure and Table Package
\usepackage{color}
\usepackage{amsfonts}
%\usepackage{ulem}    % underline/strikethrough/wave lines
\usepackage{graphicx}  % figures
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{shapes,shadows,calc}
\usepgflibrary{arrows}

@@

\let\springervec\vec
\let\vec\origvec
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

@@

\usepackage{palatino}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\usepackage{xcolor}
\usetikzlibrary{shapes,arrows,chains}

@@

\section{Use-case: The BIP Failure-Timer architecture}
\TODO{Describe here the full Failure-Timer architecture, as taken from the Avocs paper, including an informal presentation of the original BIP system, its full translation into a pNet system (pLTSs and pNet nodes), and the generated Open Automaton}
\NOTE{All figures are already available from the Avocs paper sources.}

@@

\subsubsection{Definition of pLTS}
We define pLTS as a triple: $pLTS\triangleq\ll S, s_{0}, \rightarrow \gg$ where:
\begin{itemize}
	\item $S$ is a set of states, and $s_{0}\in S$ is the initial state.
	\item $\rightarrow \subseteq S\times L\times S$ is the transition relation, with $L$ the set of labels of form $<\alpha, e_{b}, (x_{j} := e_{j})^{(j\in J)}>$, where $\alpha\in A_{V}$ is a parameterised action, $e_{b}\in B_{V}$ is a guard, and expression $E_{P}\cup A_{P}$ are assigned to $x_{j}$. If $s\xrightarrow{<\alpha, e_{b}, (x_{j}:=e_{j})^{j\in J}>}s' \in \rightarrow$, then $vars(e_{b})\subseteq vars(s)\cup vars(\alpha)$, and $\forall j\in J. vars(e_{j})\subseteq vars(s)\wedge x_{j}\in vars(s')$.
\end{itemize}

@@

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, or a receive variable $?x$. An action $\alpha$ is a parameterised action $c(arg_{i})$ or a silent action $\tau$. $p$ is an atomic proposition and $X$ is a set of variables. $\neg$, $\vee$ and $\wedge$ are respectively negation, disjunction and conjunction operator on action formula and state formula. $[\alpha]\varphi$ expresses "whenever an action $\alpha$ happens, we will do $\varphi$", and $\langle\alpha\rangle$ expresses "an action $\alpha$ takes effective, we will do $\varphi$". $x_{i}\in X$ is a free variable bound in state formula $\varphi$. $\mu$ and $\nu$ are respectively the minimum fixed point and the maximum fixed point.\TODO{Tengfei: Must more explaination}

We employ the case study from timer component\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by pM$\mu$ formula:

@@

At first, we can do $\langle start?M\rangle$ and then it enters a loop. After that, either we can do a $timeout$, and then it is done and the result is true. Or we can do a $tick$ and then we continue the loop with decrementing the value $t$ by one.

@@

\TODO{Tengfei: Be more precise}

@@

A formula is said to be alternation-free, when the alternation depth is one, without mutually recursive greatest and least fixed-point operators. A formula expressed with pM$\mu$ is alternation-free. But the formula with first-order $\mu$-calculus have to deal with alternation depth.\TODO{definition of alternation depth]}
\NOTE{A formula expressed with pM$\mu$ is alternation-free. !!!NO}

@@

\subsection{The difference between pLTS and open automata}

@@

		\If{B=VALUE(b)}
		\State pop(); 
		\If{s.status = VISITED($b'$) and $b'\neq b$}
		\State s.status = VISITED($b$); 
		\EndIf
		\EndIf
	\end{algorithmic} 
	\caption{$close(s)$}
	\label{alg:algorithm1}
\end{algorithm}

@@

	\end{algorithmic} 
	\caption{check(s)}
	\label{alg:algorithm2}
\end{algorithm}

\begin{thebibliography}{4}

>>>>

%******Figure and Table Package
\usepackage{color}
\usepackage{amsfonts}
%\usepackage{ulem}    % underline/strikethrough/wave lines
\usepackage{graphicx}  % figures
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{figlatex}
\usepackage{graphics}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{shapes,shadows,calc}
\usepgflibrary{arrows}

@@

\let\springervec\vec
\let\vec\origvec
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\setcounter{tocdepth}{3}

@@

\usepackage{palatino}
\usetikzlibrary{shapes.geometric, arrows}
\usepackage{xcolor}
\usetikzlibrary{shapes,arrows,chains}

@@

\section{Use-case: The BIP Failure-Timer architecture}

\TODO{Describe here the full Failure-Timer architecture, as taken from the Avocs paper, including an informal presentation of the original BIP system, its full translation into a pNet system (pLTSs and pNet nodes), and the generated Open Automaton}
\NOTE{All figures are already available from the Avocs paper sources.}

The BIP architecture includes two kinds of components, coordinator component and operand component. The operand component just has some ports used to communicate with other component. While the coordinator components don't only have some ports, but they include a parameterised labelling transition system (pLTS) in them. In a pLTS, some nodes and edges between these nodes are built to describe the coordinator component. In a pLTS, a node denotes the state of the component, while a edge means the transition from one state to another. Specifically, a port in a coordinator component is used to label a transition, and the number of transitions is the same as that of ports.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/BIPspec-ArchFailureTimerMax-v3.pdf}
  \caption{The BIP specification of the Failure Monitor architecture}
  \label{schema:ArchFailure:BIP}
\end{figure}

Fig.~\ref{schema:ArchFailure:BIP} shows the BIP specification of the Failure-Timer architecture from~\cite{xudong2017smt}.  $Timer$ and $Controller$ are the coordinator components, and $B$ is an operand component. $Timer$ has four ports, like $start$, $tick$, $resume$ and $timeout$, and these ports are labelling the transitions between the two states $t_{0}$, and $t_{1}$. And $Ctrl$ has four ports, like $fail$, $reset$, $resume$ and $timeout$, and these ports are labelling the transitions between the three states $s_{0}$, $s_{1}$ and $s_{2}$. The operand component $B$ has three ports $resume$, $fail$ and $finish$.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\columnwidth]{figures/FailureTimerOA.pdf}
  \caption{The Open Automaton of the Failure Monitor architecture}
  \label{schema:ArchFailure:OA}
\end{figure}

In~\cite{xudong2017smt}, an approach is present to unfold the open pNet to open automata. Fig.~\ref{schema:ArchFailure:OA} shows the open automaton unfolded from the Failure-Timer architecture in Fig~\ref{schema:ArchFailure:BIP}. The state is the composition of the states of each pLTS in the coordinator component, while the transition is labelled with a composition of the corresponding action in each component. For example, in the BIP specification, $Timer.start$ synchronises with $Ctrl.fail$, because there is a connector between the two ports. When occurrence of the transition $s_{0}\xrightarrow{fail}s_{1}$ will trigger the execution of the transition $t_{0}\xrightarrow{start, t:=0}t_{1}$. At the same time, the port $Ctrl.fail$ will synchronise with $B.fail$. The connected ports in Failure-Timer architecture form a synchronisation vector. The synchronisation will return to a special action, which can be used to label a transition.

@@

\subsubsection{Definition of pLTS}
We define pLTS as a triple: $pLTS\triangleq\ll S, s_{0}, \rightarrow \gg$ where:
\begin{itemize}
	\item $S$ is a set of states, and $s_{0}\in S$ is the initial state.
	\item $\rightarrow \subseteq S\times L\times S$ is the transition relation, with $L$ the set of labels of form $\langle\alpha, e_{b}, (x_{j} := e_{j})^{(j\in J)}\rangle$, where $\alpha\in A_{V}$ is a parameterised action, $e_{b}\in B_{V}$ is a guard, and expression $E_{P}\cup A_{P}$ are assigned to $x_{j}$. If $s\xrightarrow{\langle\alpha, e_{b}, (x_{j}:=e_{j})^{j\in J}\rangle}s' \in \rightarrow$, then $vars(e_{b})\subseteq vars(s)\cup vars(\alpha)$, and $\forall j\in J. vars(e_{j})\subseteq vars(s)\wedge x_{j}\in vars(s')$.
\end{itemize}

@@

 where, $c$ is an action name expressing the communication port. $arg_{i}$ is a list of arguments, and each $arg_{i}$ is either a send expression $!e$, or a receive variable $?x$. An action $\alpha$ is a parameterised action $c(arg_{i})$ or a silent action $\tau$. $p$ is an atomic proposition and $X$ is a set of variables. $\neg$, $\vee$ and $\wedge$ are respectively negation, disjunction and conjunction operator on action formula and state formula. $[\alpha]\varphi$ expresses "whenever an action $\alpha$ happens, we will do $\varphi$", and $\langle\alpha\rangle$ expresses "an action $\alpha$ takes effective, we will do $\varphi$". $x_{i}\in X$ is a free variable bound in state formula $\varphi$. $\mu$ and $\nu$ are respectively the minimum fixed point and the maximum fixed point.\TODO{Tengfei: Must more explaination}\NOTE{Done}

We employ the case study from timer component\cite{mavridou2016architectrue}. A liveness property defines that good thing finally happens. In this case study, intuitively the timer finally returns to the initial state after some execution steps. The property can be expressed by a pM$\mu$ formula:

@@

\ERIC{this description is too operational (like a state machine), I would like a more logical description of the formula. One way is to use the vocabulary of ``Specification patterns'', like ``it is possible'', ``it is mandatory'', ``do something until some condition'', etc}
At first, we can do $\langle start?M\rangle$ and then it is mandatory to enters a loop with an assignment $t := M$. In the loop, it is possible for us to do a $timeout$ until  $t$ is equal to 0, and then it is done and the result is true. Or we can do a $tick$ and then we continue the loop with decrementing the value $t$ by one.

@@

\TODO{Tengfei: Be more precise}\NOTE{Done}

@@

A formula is said to be alternation-free, when the alternation depth is one, without mutually recursive greatest and least fixed-point operators. In the Timer example, the alternation depth is 1. \TODO{definition of alternation depth]}

@@

\subsection{The difference between pLTS and open automata}
 pLTS is a labelling transition system with parameterised action. Open pNet is a composition of some pLTSs with synchronisation vector and some holes. A BIP architecture of an open pNet includes coordinator components and some operand components. An open automaton is formed by the unfolding of an open pNet.

@@

		\If{B=VALUE(b)}
		\State pop(); 
		\If{s.status = VISITED($b'$) and $b'\neq b$}
		\State s.status = VISITED($b$); 
		\EndIf
		\EndIf
	\end{algorithmic} 
	\caption{$close(s)$}
	\label{alg:algorithm2}
\end{algorithm}

@@

	\end{algorithmic} 
	\caption{check(s)}
	\label{alg:algorithm3}
\end{algorithm}

\begin{thebibliography}{4}
	\bibitem{xudong2017smt}Xudong Qin, Simon Bliudze, Eric Madelaine, Min Zhang. Using SMT engine to generate Symbolic Automata -Extended version. [Research Report] RR-9177, Inria
